The Sub-Millimetre User Reduction Facility: an iterative map-maker for SCUBA-2

### 2012-06-01 smurf_paper.tex
first stab at extended example m17

One simple test of the calculated noise properties is to compare the
signal and jackknife SNR distributions with ideal Gaussians. The top
panel of Fig.~\ref{fig:lockman_hist} shows the whitened (but not match
filtered) signal (blue) and jackknife (histograms), along with a
Gaussian (mean 0, standard deviation 1, and area normalized to the
number of map pixels) as a dashed line. In this case, it is clear that
the SNR distributions for both maps are nearly indistinguishable from
the theoretical distribution of white noise. This result shows us
that: (i) the whitening filter appears to have removed correlated
large-scale noise since the jackknife map histogram is consistent with
white noise; and (ii) any potential astronomical signals are small
compared to the typical white noise in most map pixels (unsurprising
given the appearance of Fig.~\ref{fig:lockman_maps}b). Next, we
examine the SNR histograms for maps processed with the matched filter
in the bottom panel of Fig.~\ref{fig:lockman_hist}. Again, the
histogram of the jackknife SNR data appears consistent with pure
noise. However, the signal map now deviates significantly from a
Gaussian distribution, with a clear \emph{positive} tail (as one would
expect for emitting sources). In fact, integrating the positive tails
beyond our 3.8-$\sigma$ source-detection threshold (vertical solid
line) yields 229 map pixels in the signal map, compared with 3 in the
jackknife map (out of 80603 pixels in the entire region). Note that
given the small pixel size of the map, several pixels generally
contribute to each peak.

@@

As an additional external check, we have over-plotted orange ``$+$''
signs at the locations of 1.4\,GHz radio sources from \citet{owen2008}
with $S\gsim15$\,$\mu$Jy. Such radio catalogues have historically
proven invaluable for the precise identifications of high-redshift
submillimetre galaxies due to their relatively low surface densities
(compare with optical catalogues, for example), and a strong
correlation between the radio and submillimetre emission mechanisms
\citep[e.g.][]{smail2000,pope2006,ivison2007,chapin2009b}. Taking a
representative search radius of 8\,arcsec from these studies with
similar SNR sources and source sizes (the same size as the blue
circles), 9 out of 10 peaks in the smoothed signal map have potential
radio counterparts, whereas the single peak in the smoothed jackknife
map does not lie near any radio source.

Note that for significantly deeper future SCUBA-2 maps, in which the
RMS in a PSF-smoothed map is dominated by point-source confusion,
rather than instrumental noise, a modified matched-filter will offer
improved results. See Appendix~A in \citet{chapin2011}, which shows
how to include confusion (when known \emph{a priori}) explicitly as a
noise term in the calculation of such filters.

\subsection{Bright extended emission}
The hardest thing to do with \scuba\ at the moment. Try to generate a
source/background mask to assist the map solution based on thresholds
in \snr. Compare with BOLOCAM stuff.

@@

default: converges after 21 iterations

>>>>

One simple test of the calculated noise properties is to compare the
signal and jackknife SNR distributions with ideal Gaussians. The top
panel of Fig.~\ref{fig:lockman_hist} shows the whitened (but not match
filtered) signal (blue) and jackknife (histograms), along with a
Gaussian (mean 0, standard deviation 1, and area normalized to the
number of map pixels) as a dashed line. In this case, it is clear that
the SNR distributions for both maps are nearly indistinguishable from
the theoretical distribution of white noise. This result shows us
that: (i) the whitening filter appears to have removed correlated
large-scale noise since the jackknife map histogram is consistent with
white noise; and (ii) any potential astronomical signals are small
compared to the typical white noise in most map pixels (unsurprising
given the appearance of Fig.~\ref{fig:lockman_maps}b). Next, we
examine the SNR histograms for maps processed with the matched filter
in the bottom panel of Fig.~\ref{fig:lockman_hist}. Again, the
histogram of the jackknife SNR data appears consistent with pure
noise. However, the signal map now deviates significantly from a
Gaussian distribution, with a clear \emph{positive} tail (as one would
expect for emitting sources). In fact, integrating the positive tails
beyond our 3.8-$\sigma$ source-detection threshold (vertical solid
line) yields 229 map pixels in the signal map, compared with 3 in the
jackknife map (out of 80603 pixels in the entire region). Note that
given the small pixel size of the map, several pixels generally
contribute to each peak. Note that this procedure gives only a flavour
of the analysis that is usually required to produce a robust source
lists. Additional tests, along with a careful consideration of
completeness and false-positive rates usual require a series of Monte
Carlo simulations that are beyond the scope of this work. We direct
the interested reader to a selection of papers on the subject:
\citet{scott2002,coppin2006,perera2008,weiss2009} and
\citet{chapin2011}.

@@

As an additional external check, we have over-plotted orange ``$+$''
signs at the locations of 1.4\,GHz radio sources from \citet{owen2008}
with $S\gsim15$\,$\mu$Jy. Such radio catalogues have historically
proven invaluable for the precise identifications of high-redshift
submillimetre galaxies due to their relatively low surface densities
(compare with optical catalogues, for example), and a strong
correlation between the radio and submillimetre emission mechanisms
\citep[e.g.][]{smail2000,pope2006,ivison2007,chapin2009b}. Taking a
representative search radius of 8\,arcsec from these studies with
similar SNR sources and source sizes (the same size as the blue
circles), 9 out of 10 peaks in the smoothed signal map have potential
radio counterparts, whereas the single peak in the smoothed jackknife
map does not lie near any radio source. Again, a proper
cross-identification analysis must inevitably include simulations to
establish completeness, false-positive rates, as well as the effects
of point source clustering and confusion. See the aforementioned
papers and references therein for examples.

For future, significantly deeper SCUBA-2 maps, in which the RMS in a
PSF-smoothed map is dominated by point-source confusion, rather than
instrumental noise, a modified matched-filter will offer improved
results. See Appendix~A in \citet{chapin2011}, which shows how to
include confusion (when known \emph{a priori}) explicitly as a noise
term in the calculation of such filters.

\subsection{Bright extended emission}

In this final example, we analyze a map of M17 which contains bright,
extended emission, in contrast with the earlier two examples. A
promise of \scuba\ is to recover larger angular scales than what was
possible with earlier-generation ground-based submillimetre
cameras. The two main factors contributing to this expectation are:
(i) a larger array footprint on the sky which should make the
subtraction of common sky-emission signals more effective; and (ii)
fast detectors (with a correspondingly fast sample rate) to enable
higher slew-rates (up to 600\,arcsec\,s$^{-1}$) to capture as much
large-scale structure as possible above the detector $1/f$ knees. We
produce maps of observation 11 on 20110531 using the 850\,\micron\
array. It is a rotating PONG convering a diameter of 0.375\,deg, with
a scan speed of 300\,arcsec\,s$^{-1}$, and a transverse spacing of
180\,arcsec, taking 37.5\,min to complete.

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{m17.pdf}
\caption{An 850\,\micron\ rotating pong map of M17. Intensity is
logarithmically scaled between -0.0003\,pw (white) and +0.01\,pW
(black). Iteration numbers are given in the corner of each
panel. Panels (a) and (b) show the results for a reduction using the
default parameters (the solution halted after reaching the default
map-based convergence criterion in 30 iterations). Panel (a) also
depicts the array footprint (position angle indicative of the start of
the observation), and a 300\,arcsec line shows the spatial scale
corresponding to the default \model{FLT} high-pass filter. Similar to
Fig.~\ref{fig:pointmaps}c, the lack of prior constraints on the map
leads to degeneracies, and therefore ripples on scales larger than the
array footprint (due to common-mode subtraction) and filter scale
(especially evident around the brightest peak at the centre of the
map). Panels (c) and (d) show the ``bright extended'' reduction, in
which a zero-mask is created iteratively from all of the pixels that
exceed a SNR of 5. While this region (red contour) only encompasses
the brightest peaks early in the solution, the final iteration it
identifies most of the bright, extended emission, and it significantly
helps with negative ringing.}
\label{fig:m17}
\end{figure*}

The default reduction of these data is shown in the top panels of
Fig.~\ref{fig:m17}, after 1, 30 iterations (by this last
iteration the map-based convergence criterion was achieved). The first
panel also depicts the array footprint, and the angular scale
corresponding the high-pass filter edge. Much like the reduction of a
point source without any prior constraints
(Fig.~\ref{fig:pointmaps}c), degeneracies on scales larger than the
array footprint and filter produce ripples in the map.

Unlike the case of a known point-source (Section~\ref{sec:point}), it
may not be possible for the observer to define, in advance, a mask to
define regions of source and blank sky. Indeed, for this map, much of
the field clearly contains extended structures. Furthermore, the goal
of such maps may be to detect \emph{previously unknown} cool, dense
regions of the ISM that may not have appeared at other wavelengths
(e.g. the first optically-thick cloud-collapse stages of
star-formation). While the option does exist for the user to supply
their own mask, a facility has been added to SMURF to generate one
automatically by flagging pixels above some SNR threshold after each
iteration and is referred to as the ``bright extended'' configuration.

The results of this automatic masking are shown in the bottom panels
of Fig.~\ref{fig:m17}. After the first iteration, only narrow regions
around the brightest peaks are identified. However, as the solution
progresses, the negative bowls around the bright sources are slowly
reduced and the mask ``grows'' out from the brightest areas. By the
final iteration, most of the obvious structures in the data are
encompasses by the mask.

Note that alternatives to this approach do exist for other iterative
map-makers. For example, \citet{kovacs2008} restricts the solution to
a small number of iterations to avoid divergence on large-scales. For
BOLOCAM data of the Galactic Plane, \citet{aguirre2010} used a
Maximum-Entropy filtering step to suppress large-scales. The
zero-masking implemented in SMURF is similar in principle to the
\citet{aguirre2010}, while being slightly conceptually
simpler. However, since both methods are non-linear, detailed
simulations would be required to determine the SNR-dependent transfer
function.

While the reduction in the bottom panels of Fig.~\ref{fig:m17} is
clearly superior in at least a cosmetic sense to those in the top
panels, it is important to quantify both the noise properties of the
maps and their response to real structures. Similar to the previous
section, we will use a jackknife test to estimate the noise, and then
inject known sources into the data to observe how they are attenuated.

%A second map of M17 using the same observing mode was made shortly
%after the first, observation 13 on 20110511. We produce jackknife maps
%from the differences between the two ... angular power spectrum of the
%noise (should be worse at large-scale for the default map).

Next, we inject simulated clouds into a relative empty portion of the
observation 11 M17 map. It is created by filtering white noise first
with a power-law index $alpha=$bla (a reasonable approximation for
Galactic Cirrus... cite bla bla), and then with a 14.5\,arcsec
Gaussian to model the effect of the SCUBA-2 optical response. Finally,
the signal is apodized ... Compare power-spectrum before and after.

One fairly arbitrary parameter in our data reduction is the choice of
angular scale for the \model{FLT} filter. The current default of
300\,arcsec\ was chosen ``by-eye'' as a compromise that would provides
aesthetically reasonable results for a range of observing
modes. However, there is a simple tradeoff that warrants
experimentation for specific fields. By choosing a larger scale (lower
frequency in the bolometers signals) for the filter, potentially more
structure will be faithfully re-produced in the map (up to the scale
of the array footprint, at which point the common-mode removal will
dominate). However, lower the filter edge can potentially let through
more independent low-frequency noise. We have therefore re-run the
previous tests using two other scales for the filters, and show the
resulting maps and angular SNR spectra in Figs ... .

@@
